### 2021-10-28 reviews/01_ELE/01_draft_revised.tex
materials

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0003-0193-5441}{Salomé\,Bouskila}%
%
\,\textsuperscript{1,‡}\quad %
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,3,2}\quad %
\href{https://orcid.org/0000-0003-4036-977X}{Ceres\,Barros}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-2151-6693}{Dominique\,Caron}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0003-0452-6993}{Maxwell J\,Farrell}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0002-9935-1366}{Marie-Josée\,Fortin}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0003-3220-6161}{Victoria\,Hemming}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-4104-9463}{Benjamin\,Mercier}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0001-5785-8321}{Rogini\,Runghen}%
%
\,\textsuperscript{7}\quad %
\href{https://orcid.org/0000-0002-3454-0633}{Giulio V.\,Dalla Riva}%
%
\,\textsuperscript{8}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Québec Centre for
Biodiversity Sciences, Montréal,
Canada\quad \textsuperscript{3}\,Université de Sherbrooke, Sherbrooke,
Canada\quad \textsuperscript{4}\,Department of Forest Resources
Management, University of British Columbia, Vancouver,
Canada\quad \textsuperscript{5}\,Department of Biology, McGill
University, Montréal, Canada\quad \textsuperscript{6}\,Department of
Ecology \& Evolutionary Biology, University of Toronto, Toronto,
Canada\quad \textsuperscript{7}\,Centre for Integrative Ecology, School
of Biological Sciences, University of Canterbury, Canterbury, New
Zealand\quad \textsuperscript{8}\,School of Mathematics and Statistics,
University of Canterbury, Canterbury, New Zealand

@@

\vfill
Despite their importance in many ecological processes, collecting data
and information on ecological interactions is an exceedingly challenging
task. For this reason, large parts of the world have a data deficit when
it comes to species interactions, and how the resulting networks are
structured. As data collection alone is unlikely to be sufficient,
community ecologists must adopt predictive methods. Here we develop such
a method, relying on graph embedding and transfer learning to assemble a
predicted list of trophic interactions between Canadian mammals. This
interaction list is derived from the European food web, despite sharing
only 5\% of common species with Canada. The results of the predictive
model are compared against databases of recorded pairwise interactions,
showing that we correctly recover over 95\% of known interactions. We
provide guidance on how this method can be adapted by substituting some
approaches or predictors in order to make it more generally applicable.

@@

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen
2016). First, networks within a location are difficult to sample
properly (Jordano 2016a, b), resulting in a widespread ``Eltonian
shortfall'' (Hortal \emph{et al.} 2015), \emph{i.e.} a lack of knowledge
about inter and intra specific relationships. This first challenge has
been, in large part, addressed by the recent emergence of a suite of
methods aiming to predict interactions within \emph{existing} networks,
many of which are reviewed in Strydom \emph{et al.} (2021a). Second,
recent analyses based on collected data (Poisot \emph{et al.} 2021a) or
metadata (Cameron \emph{et al.} 2019) highlight that ecological networks
are currently studied in a biased subset of space and bioclimates, which
impedes our ability to generalize any local understanding of network
structure. Meaning that, although the framework to address
incompleteness \emph{within} networks exists, there would still be
regions for which, due to a \emph{lack} of local interaction data, we
are unable to infer potential species interactions. Having a general
solution for inferring \emph{plausible} interactions (despite the
unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom \emph{et al.}
(2021a) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

Here, we present a general method to recommend possible trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. There is a plurality of
measures of species similarities that can be used for metaweb
reconstruction (see \emph{e.g.} Morales-Castilla \emph{et al.} 2015);
however, phylogenetic proximity has several desirable properties when
working at large scales. Gerhold \emph{et al.} (2015) made the point
that phylogenetic signal captures diversification of characters (large
macro-evolutionary process), but not necessarily community assembly
(fine ecological process); Dormann \emph{et al.} (2010) previously found
very similar conclusions. Interactions tend reflect a phylogenetic
signal because they have a conserved pattern of evolutionary convergence
that encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares \emph{et al.} 2009; Mouquet \emph{et al.} 2012), and -
most importantly - retain this signal even when it is not detectable at
the community scale (Hutchinson \emph{et al.} 2017; Poisot \& Stouffer
2018). Finally, species interactions at macro-ecological scales seem to
respond mostly to macro-evolutionary processes (Price 2003); which is
evidenced by the presence of conserved backbones in food webs (Dalla
Riva \& Stouffer 2016; Mora \emph{et al.} 2018), strong evolutionary
signature on prey choice (Stouffer \emph{et al.} 2012), and strong
phylogenetic signature in food web intervality (Eklöf \& Stouffer 2016).
Phylogenetic reconstruction has also previously been used within the
context of ecological networks, namely understanding ancestral
plant-insect interactions (Braga \emph{et al.} 2021). Taken together,
these considerations suggest that phylogenies can reliably be used to
transfer knowledge on species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (vulnerability traits representing species at
the lower trophic-level and generality traits representing species at
higher trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano \emph{et
al.} 2020b, a) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham \emph{et al.} 2019),
infer a metaweb for the Canadian mammalian species pool (interactions
are treated as unknown in this instance). Following the definition of
Dunne (2006), a metaweb is a network analogue to the concept of a
regional species pool. Specifically, a metaweb is an inventory of all
\emph{possible} interactions within species likely to occurr within a
spatially delimited area (the network \(\gamma\)-diversity, in a sense).
The metaweb is, therefore, \emph{not} a prediction of the food web at
any specific locale within the frontiers of the spatial area it
recovers, and will in fact have a different structure (notably by having
a larger connectance; see \emph{e.g.} Wood \emph{et al.} 2015). These
local food webs are expected to be a subset of both the species and the
interactions of their metaweb, and have been called ``metaweb
realizations'' (Poisot \emph{et al.} 2015). The difference between a
food web at a specific location and the metaweb has to do with a variety
of mechanisms, including species co-occurrence, local environmental
conditions, and local distribution of functional traits. Nevertheless,
the metaweb represents the total of functional, phylogenetic, and
macroecological processes (Morales-Castilla \emph{et al.} 2015), and
therefore still holds valuable ecological information. Because the
metaweb can be down-sampled given appropriate knowledge of local species
composition (the network \(\alpha\)-diversity, in a sense), it is
possible to infer what may drive the structure of food webs at finer
spatial scales. This has been done for example for
tree-gallers-parasitoid systems (Gravel \emph{et al.} 2018), fish
trophic interactions (Albouy \emph{et al.} 2019), tetrapods trophic
interactions (O'Connor \emph{et al.} 2020), and crop-pests networks
(Grünig \emph{et al.} 2020). Whereas the original metaweb definition was
based on presence/absence, we focus on \emph{probabilistic} metawebs;
not only does our method recommed interactions that may exist, it gives
each interaction a score that is mathematically equivalent to the chance
of success of a Bernoulli trial (see \emph{e.g.} Poisot \emph{et al.}
2016), which allows properly weigh interactions as a function of how
likely they are.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the result, when validated against
known but fracational data of trophic interactions between Canadian
mammals, our model achieves a predictive accuracy of approx. 91\%. It
should be reiterated that the framework presented in
fig.~\ref{fig:concept} is amenable to changes; notably, the measure of
similarity may not be phylogeny, and can be replaced by information on
foraging (Beckerman \emph{et al.} 2006), cell-level mechanisms
(Boeckaerts \emph{et al.} 2021), or a combination of traits and
phylogenetic structure (Stock 2021). Most importantly, although we focus
on a trophic system, it is an established fact that different types of
interactions do themselves interact (Kéfi \emph{et al.} 2012; see
\emph{e.g.} Kawatsu \emph{et al.} 2021). Future development of metaweb
inference techniques should cover the prediction of multiple interaction
types.

@@

We use data from the European metaweb assembled by Maiorano \emph{et
al.} (2020b).

\textbf{TANYA TODO} description of the data

We induced the subgraph corresponding to all mammals by matching species
names in the original network to the GBIF taxonomic backbone (GBIF
Secretariat 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié \emph{et al.} 2021)). All
nodes had valid matches to GBIF at this step, and so this backbone is
used for all name reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer. We used the mammalian consensus supertree by
Upham \emph{et al.} (2019), for which all approximatively 6000 names
have been similarly matched to their GBIF valid names. This step allows
us to place each node of the mammalian European metaweb in the
phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by Maiorano \emph{et al.} (2020b) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko \emph{et al.} 2011) to generate latent traits representing
a low-dimensional embedding of the network; these traits give an
unbiased estimate of the node's position in the latent feature spaces.
Then, we map these latent traits onto a reference phylogeny (other
distance-based measures of species proximity that allow for the
inference of features in the latent space can be used, for example the
dissimilarity in functional traits). Based on the reconstructed latent
traits for species in the destination species pool, a Random Dot Product
Graph model (hereafter RDPG; Young \& Scheinerman 2007) predicts the
interaction between species through a function of the nodes' features
through matrix multiplication. Thus, from latent traits and node
position, we can infer interactions.

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
Holm (2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information that we may thought; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy \emph{et al.} 2009); in fact, in an instance of ecological
forecasting of spatio-temporal systems, model-free approaches
(\emph{i.e.} drawing all of their information from the data)
outperformed model-informed ones (Perretti \emph{et al.} 2013).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically close
predators should share phylogenetically close preys. In addition, recent
advances show that the latent variables produced this way can be used to
predict \emph{de novo} network edges. Interstingly, the latent variables
do not need to be prouced by decomposing the network itself; in a recent
contribution, Runghen \emph{et al.} (2021) show that deep artificial
neural networks are able to reconstruct the left and right subspaces of
an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom
\emph{et al.} 2021b) and to allow highly accurate prediction of the
interactions within a single network (Poisot \emph{et al.} 2021b). Under
SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.
As an aside, most ecologists are indirectly familiar with SVD: Principal
Component Analysis is a special case of SVD, which is more sensitive to
numerical instabilities (see notably Shlens 2014).

@@

It must be noted that despite bringing in a large amount of information
from the European species pool and interactions, the Canadian metaweb
has distinct structural properties. Following an approach similar to
Vermaat \emph{et al.} (2009), we show in Supp. Mat. 3 that not only can
we observed differences in a multivariate space between the European and
Canadian metaweb, we can also observe differences in the same space
between random subgraphs from these networks. These results line up with
the studies spatializing metawebs that have been discussed in the
introduction: changes in the species pool are driving local structural
changes in the networks.

>>>>

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0003-0193-5441}{Salomé\,Bouskila}%
%
\,\textsuperscript{1,‡}\quad %
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,3,2}\quad %
\href{https://orcid.org/0000-0003-4036-977X}{Ceres\,Barros}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-2151-6693}{Dominique\,Caron}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0003-0452-6993}{Maxwell J\,Farrell}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0002-9935-1366}{Marie-Josée\,Fortin}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0003-3220-6161}{Victoria\,Hemming}%
%
\,\textsuperscript{7}\quad %
\href{https://orcid.org/0000-0002-4104-9463}{Benjamin\,Mercier}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0001-5785-8321}{Rogini\,Runghen}%
%
\,\textsuperscript{8}\quad %
\href{https://orcid.org/0000-0002-3454-0633}{Giulio V.\,Dalla Riva}%
%
\,\textsuperscript{9}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Quebec Centre for
Biodiversity Science, Montréal,
Canada\quad \textsuperscript{3}\,Département de Biologie, Université de
Sherbrooke, Sherbrooke, Canada\quad \textsuperscript{4}\,Department of
Forest Resources Management, University of British Columbia, Vancouver,
B.C., Canada\quad \textsuperscript{5}\,Department of Biology, McGill
University, Montréal, Canada\quad \textsuperscript{6}\,Department of
Ecology \& Evolutionary Biology, University of Toronto, Toronto,
Canada\quad \textsuperscript{7}\,Department of Forest and Conservation
Sciences, University of British Columbia, Vancouver,
Canada\quad \textsuperscript{8}\,Centre for Integrative Ecology, School
of Biological Sciences, University of Canterbury, Canterbury, New
Zealand\quad \textsuperscript{9}\,School of Mathematics and Statistics,
University of Canterbury, Canterbury, New Zealand

@@

\vfill
Despite their importance in many ecological processes, collecting data
and information on ecological interactions is an exceedingly challenging
task. For this reason, large parts of the world have a data deficit when
it comes to species interactions, and how the resulting networks are
structured. As data collection alone is unlikely to be sufficient,
community ecologists must adopt predictive methods. Here, we develop
such a method relying on graph embedding and transfer learning to
assemble a predicted list of trophic interactions between Canadian
mammals. This interaction list is derived from the European food web,
despite sharing 4\% of common species with Canada. The results of the
predictive model are compared against databases of recorded pairwise
interactions, showing that we correctly recover 91\% of known
interactions. We provide guidance on how this method can be adapted by
substituting some approaches or predictors in order to make it more
generally applicable.

@@

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen
2016). First, ecological networks within a location are difficult to
sample properly (Jordano 2016a, b), resulting in a widespread ``Eltonian
shortfall'' (Hortal \emph{et al.} 2015), \emph{i.e.} a lack of knowledge
about inter- and intra- specific relationships. This first challenge has
been, in large part, addressed by the recent emergence of a suite of
methods aiming to predict interactions within \emph{existing} networks,
many of which are reviewed in Strydom \emph{et al.} (2021a). Second,
recent analyses based on collected data (Poisot \emph{et al.} 2021a) or
metadata (Cameron \emph{et al.} 2019) highlight that ecological networks
are currently studied in a biased subset of space and bioclimates, which
impedes our ability to generalize any local understanding of network
structure. Meaning that, although the framework to address
incompleteness \emph{within} networks exists, there would still be
regions for which, due to a \emph{lack} of local interaction data, we
are unable to infer potential species interactions. Having a general
solution for inferring \emph{potential} interactions (despite the
unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom \emph{et al.}
(2021a) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

Here, we present a general method to recommend potential trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. There is a plurality of
measures of species similarities that can be used for inferring
\emph{potential} species interactions \emph{i.e.} metaweb reconstruction
(see \emph{e.g.} Morales-Castilla \emph{et al.} 2015); however,
phylogenetic proximity has several desirable properties when working at
large scales. Gerhold \emph{et al.} (2015) made the point that
phylogenetic signal captures diversification of characters (large
macro-evolutionary process), but not necessarily community assembly
(fine ecological process); Dormann \emph{et al.} (2010) previously found
very similar conclusions. Interactions tend to reflect a phylogenetic
signal because they have a conserved pattern of evolutionary convergence
that encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares \emph{et al.} 2009; Mouquet \emph{et al.} 2012), and -
most importantly - retain this signal even when it is not detectable at
the community scale (Hutchinson \emph{et al.} 2017; Poisot \& Stouffer
2018). Finally, species interactions at macro-ecological scales seem to
respond mostly to macro-evolutionary processes (Price 2003); which is
evidenced by the presence of conserved backbones in food webs (Dalla
Riva \& Stouffer 2016; Mora \emph{et al.} 2018), strong evolutionary
signature on prey choice (Stouffer \emph{et al.} 2012), and strong
phylogenetic signature in food web intervality (Eklöf \& Stouffer 2016).
Phylogenetic reconstruction has also previously been used within the
context of ecological networks, namely understanding ancestral
plant-insect interactions (Braga \emph{et al.} 2021). Taken together,
these considerations suggest that phylogenies can reliably be used to
transfer knowledge on species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (latent vulnerability traits are more
representative of species at the lower trophic-level and latent
generality traits are more representative of species at higher
trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano \emph{et
al.} 2020b, a) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham \emph{et al.} 2019),
infer a metaweb for the Canadian mammalian species pool (interactions
are treated as unknown in this instance). Following the definition of
Dunne (2006), a metaweb is a network analogue to the regional species
pool; specifically, it is an inventory of all \emph{potential}
interactions between species from a spatially delimited area (and so
captures the \(\gamma\) diversity of interactions). The metaweb is,
therefore, \emph{not} a prediction of the food web at a specific locale
within the spatial area it covers, and will have a different structure
(notably by having a larger connectance; see \emph{e.g.} Wood \emph{et
al.} 2015). These local food webs (which captures the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
interactions, and have been called ``metaweb realizations'' (Poisot
\emph{et al.} 2015). Differences between local food web and their
metaweb are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla \emph{et
al.} 2015), it holds valuable ecological information. Specifically, it
is the ``upper bounds'' on what the composition of the local networks
can be (see \emph{e.g.} McLeod \emph{et al.} 2021). These local
networks, in turn, can be reconstructed given appropriate knowledge of
local species composition, providing information on structure of food
webs at finer spatial scales. This has been done for example for
tree-galler-parasitoid systems (Gravel \emph{et al.} 2018), fish trophic
interactions (Albouy \emph{et al.} 2019), tetrapod trophic interactions
(O'Connor \emph{et al.} 2020), and crop-pest networks (Grünig \emph{et
al.} 2020). Whereas the original metaweb definition, and indeed most
past uses of metawebs, was based on the presence/absence of
interactions, we focus on \emph{probabilistic} metawebs where
interactions are represented as the chance of success of a Bernoulli
trial (see \emph{e.g.} Poisot \emph{et al.} 2016); therefore, not only
does our method recommend interactions that may exist, it gives each
interaction a score, allowing us to properly weigh them.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes; notably, the measure
of similarity may not be phylogeny, and can be replaced by information
on foraging (Beckerman \emph{et al.} 2006), cell-level mechanisms
(Boeckaerts \emph{et al.} 2021), or a combination of traits and
phylogenetic structure (Stock 2021). Most importantly, although we focus
on a trophic system, it is an established fact that different
(non-trophic) interactions do themselves interact with and influence the
outcome of trophic interactions (Kéfi \emph{et al.} 2012; see
\emph{e.g.} Kawatsu \emph{et al.} 2021). Future development of metaweb
inference techniques should cover the prediction of multiple interaction
types.

@@

We use data from the European metaweb assembled by Maiorano \emph{et
al.} (2020b). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
This metaweb itself is a network of binary (\emph{i.e.}
presence/absence), potential two-way interactions between species pairs.

We filtered down the European metaweb to create a subgraph corresponding
to all mammals by matching species names in the original network to the
Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF
Secretariat 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié \emph{et al.} 2021)). All
nodes had valid matches to GBIF at this step, and so this backbone is
used for all name reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer (\emph{i.e.} the transfer medium). We used the
mammalian consensus supertree by Upham \emph{et al.} (2019), for which
all approximatively 6000 names have been similarly matched to their GBIF
valid names. This step allows us to place each node of the mammalian
European metaweb in the phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the International Union for Conservation of Nature
(IUCN) checklist, and selected the terrestrial and semi-aquatic species
(this corresponds to the same selection that was applied by Maiorano
\emph{et al.} (2020b) in the European metaweb). The IUCN names were, as
previously, reconciled against GBIF to have an exact match to the
taxonomy.

@@

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko \emph{et al.} 2011) to generate latent traits representing
a low-dimensional embedding of the network. As an aside, most ecologists
are indirectly familiar with SVD: Principal Component Analysis is a
special case of SVD, which is more sensitive to numerical instabilities
(see notably Shlens 2014). The latent traits give an unbiased estimate
of the node's position in the latent feature spaces and can be mapped
onto a reference phylogeny (other distance-based measures of species
proximity that allow for the inference of features in the latent space
can be used, for example the dissimilarity in functional traits). Based
on the reconstructed latent traits for species in the destination
species pool, a Random Dot Product Graph model (hereafter RDPG; Young \&
Scheinerman 2007) predicts the interaction between species through a
function of the nodes' features through matrix multiplication. Thus,
from latent traits and node position, we can infer interactions.

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy \emph{et al.} 2009); in fact, in an instance of ecological
forecasting of spatio-temporal systems, model-free approaches
(\emph{i.e.} drawing all of their information from the data)
outperformed model-informed ones (Perretti \emph{et al.} 2013).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically similar
predators should share phylogenetically similar preys. In addition,
recent advances show that the latent variables produced this way can be
used to predict \emph{de novo} network edges. Interestingly, the latent
variables do not need to be produced by decomposing the network itself;
in a recent contribution, Runghen \emph{et al.} (2021) showed that deep
artificial neural networks are able to reconstruct the left and right
subspaces of an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional metadata as predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom
\emph{et al.} 2021b) and to allow highly accurate prediction of the
interactions within a single network (Poisot \emph{et al.} 2021b). Under
SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.

@@

It must be noted that despite bringing in a large amount of information
from the European species pool and interactions, the Canadian metaweb
has distinct structural properties. Following an approach similar to
Vermaat \emph{et al.} (2009), we show in Supp. Mat. 3 that not only can
we observe differences in a multivariate space between the European and
Canadian metaweb, we can also observe differences in the same space
between random subgraphs from these networks. These results line up with
the studies spatializing metawebs that have been discussed in the
introduction: changes in the species pool are driving local structural
changes in the networks.
